---
title: "regprojfin"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Reading libraries
```{r}
library(psych)
library(DescTools)
library(olsrr)
library(MASS)
library(ggplot2)
library(gridExtra)
```


#Reading dataset, removing variables to be predicted later
```{r}
df=read.csv("utrechthousingsmall.csv")
names(df)
#setting seed for reputability
set.seed(1145)
#selecting  2 random rows to predict later on
selected_rows <- sample(nrow(df), 2)
selected_rows
# Remove the selected rows
df2 <- df[-selected_rows, ]
attach(df2)
```



#Histogram indepnent variable
```{r}
ggplot(df2, aes(x = retailvalue)) +
  geom_histogram(binwidth = 100000, fill = "lightblue", color = "black") +
  labs(title = "Retail Value", x = "Reatil Value", y = "Frequency")
```


#Histogram cont variables 
```{r}
p1<-ggplot(df2, aes(x = house.area)) +
  geom_histogram(binwidth = 10, fill = "lightblue", color = "black") +
  labs(title = "House Area", x = "House Area", y = "Frequency")

p2<-ggplot(df2, aes(x = lot.area)) +
  geom_histogram(binwidth = 10, fill = "lightblue", color = "black") +
  labs(title = "Lot Area", x = "lot Area ", y = "Frequency")

p3<-ggplot(df2, aes(x = lot.len)) +
  geom_histogram(binwidth = 2, fill = "lightblue", color = "black") +
  labs(title = "Lot Length", x = "Lot Length", y = "Frequency")

p4<-ggplot(df2, aes(x = lot.width)) +
  geom_histogram(binwidth = 1, fill = "lightblue", color = "black") +
  labs(title = "Lot Width", x = "Lot Width", y = "Frequency")
grid.arrange(p1, p2, p3, p4, ncol = 2)
```
#remaining histogrmas
```{r}
p5<-ggplot(df2, aes(x = garden.size)) +
  geom_histogram(binwidth = 10, fill = "lightblue", color = "black") +
  labs(title = "Garden Size", x = "Garden Size", y = "Frequency")

p6<-ggplot(df2, aes(x = buildyear)) +
  geom_histogram(binwidth = 10, fill = "lightblue", color = "black") +
  labs(title = "Build year", x = "Build Year", y = "Frequency")
grid.arrange(p5, p6,ncol = 2)

```

#Plot matrix 
```{r}
dfc=subset(df2,select=c(lot.len,lot.width,house.area,garden.size,buildyear,
                        taxvalue,zipcode,balcony,bathrooms,monument,x.coor,y.coor))
pairs.panels(dfc)

```

#Declaring variables as factors
```{r}
#Declraing catgorical varables as factors 
zipcode<-as.factor(zipcode)
balcony<-as.factor(balcony)
bathrooms<-as.factor(bathrooms)
energy.eff<-as.factor(energy.eff)
#y.coor<-as.factor(y.coor)
#x.coor<-as.factor(x.coor)
monument<-as.factor(monument)
```


#Intial model
```{r}
#Initial model(All variables)

model1<-lm(formula=retailvalue~lot.len+lot.width+lot.area+ house.area+
             +garden.size+balcony+buildyear+zipcode+bathrooms+
             +taxvalue+energy.eff+monument)

summary(model1)
#High R^2 and high f value, with insignificant variables  colineartity
#Colinearty is presnt as there is a large Fvalue small t ratio, so the variables are
#insignficant but model is signficant

#Checking vif and condition index checking vif
ols_coll_diag(model1)
```
#2 model dropping tax value
```{r}
#Dropping tax value, factor loading above 0.7 part of co linear group , 
#has equal values as house area in the factor loading.IN the row with highest column index and co linear with lot area as they both have values above 0.7
model2<-lm(formula=retailvalue~lot.len+lot.width+lot.area+ house.area+
             +garden.size+balcony+buildyear+zipcode+bathrooms+
             +energy.eff+monument)
summary(model2)
#ANova to check if drooping tax value is worth it 
anova(model2,model1)
#p value =0.958,fail to reject H0 reduced model is better
ols_coll_diag(model2)

```

#3 model dropping lot area
```{r}
model3<-lm(formula=retailvalue~lot.len+lot.width+ house.area+
             +garden.size+balcony+buildyear+zipcode+bathrooms+
             +energy.eff+monument)
summary(model3)
#checking removing lot area is worth it
anova(model3,model2)
#p value is 0.2387 worth it 

#Checking if there is still multicolinearty
#Checking vif and condition index checking vif
ols_coll_diag(model3)
#No multicolinearty in vif , but there is condition index
#Highest cndition index value is between zipcode and built year,  set 14
#The second highest is between lot.len ,lot.width and garden size set 13

```

#4 model dropping lot len
```{r}
#Dropping lot len the second corelated variable from model 3
model4<-lm(formula=retailvalue~lot.width+ house.area+garden.size+
             buildyear+balcony+zipcode+bathrooms+
             +energy.eff+monument)
anova(model4,model3)
#worth it, pvalue =0.277
summary(model4)
ols_coll_diag(model4) 
#No corelation in vif, but in the condition index there is corelation
#only one zip category index is signficant, other categories are related to other variables
```

#5 model dropping zipcode or built year(Testing)
```{r}
model51<-lm(formula=retailvalue~lot.len+lot.width+ house.area+
             +garden.size+balcony+zipcode+bathrooms+
             +monument)
anova(model51,model4)
#anova said we should not drop build year
model52<-lm(formula=retailvalue~lot.len+lot.width+ house.area+
              +garden.size+balcony+buildyear+bathrooms+
              +monument)
anova(model52,model4)
#anova said no drop zip code
```

#6 model and changing zipcode categroy
```{r}
#classifying zipcode as either signficant category or other 
zipcode <- ifelse(zipcode == '3528','3528','other')
model6<-lm(formula=retailvalue~lot.width+ house.area+garden.size+
             buildyear+balcony+zipcode+bathrooms+
             +energy.eff+monument)
summary(model6)
#Garden size is insignificant
ols_coll_diag(model6)
#correlated between lot width and house area
```

#Testing droping house area or lot.width  based on condition index corelation
```{r}
#Dropping lot.width
model61<-lm(formula=retailvalue~house.area+garden.size+
             buildyear+balcony+zipcode+bathrooms++energy.eff+monument)
anova(model61,model6)
#pvalue, less than 0.05 rejecct h0, full model is better will not drop lot.width

#dropping house area
model62<-lm(formula=retailvalue~lot.width+garden.size+          buildyear+balcony+zipcode+bathrooms++energy.eff+monument)
anova(model62,model6)
#pvalue, less than 0.05 rejecct h0,full model is better will not drop house.area



```
 
#7 model, Dropping insignficant variable garden.size
```{r}
model7<-lm(formula=retailvalue~lot.width+ house.area+
             buildyear+balcony+zipcode+bathrooms+
             +energy.eff+monument)
summary(model7)
anova(model7,model6)
#pvaleu 0.5 drop value
ols_coll_diag(model7)
#no corelation in vif less than 10, condition index values above than 10 tested
#first two between house area and lot width both tested with naova and said not to drop 
#intrecept with build year
```

#plot model 7
```{r}
plot(model7)
```

#Transformation, to fix nice 
```{r}
#transformations using box cox , on cont. variables
par(mfrow=c(2,2))
bcha <- boxcox(retailvalue ~ house.area)
(lambdaha <- bcha$x[which.max(bcha$y)])

bclw <- boxcox(retailvalue ~ lot.width)
(lambdalw <- bclw$x[which.max(bclw$y)])

bcby<-boxcox(retailvalue ~ buildyear)
(lambdaby <- bcby$x[which.max(bcby$y)])
```
#Model 8 wih transformation of variables 
```{r}
model8 <- lm(formula = retailvalue ~ I(lot.width^lambdalw) + I(house.area^lambdaha) +
               I(buildyear^lambdaby) + balcony + zipcode + bathrooms +
               energy.eff+monument)
summary(model8)
plot(model8)
ols_coll_diag(model8)
#High corelation in the condition index but non in the vif 
#relation energy.eff,zipcodeother,buildyear,monumnet1,house.area

```
#Checking interaction between categorical variables 
```{r}
energy<-energy.eff
#Checking interaction terms(zipcode+balcony+bathrooms+energy.eff+monument)
par(mfrow=c(2,2))

#yes 
interaction.plot(zipcode,balcony,retailvalue)
#no
interaction.plot(zipcode,bathrooms,retailvalue)
#mabye
interaction.plot(zipcode,energy,retailvalue)
#yes 
interaction.plot(balcony,bathrooms,retailvalue)
#yes 
interaction.plot(balcony,energy,retailvalue)
#one line 
interaction.plot(bathrooms,energy,retailvalue)
#maybe 
interaction.plot(monument,energy,retailvalue)
#yes
interaction.plot(monument,bathrooms,retailvalue)
#yes
interaction.plot(monument,balcony,retailvalue)
#yes
interaction.plot(monument,zipcode,retailvalue)
```
#Testing einteraction variables that gave yes
```{r}
m11<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+balcony*zipcode+bathrooms+
          +energy.eff+monument)
summary(m11)
#interaction terms are insignificant and balcony variable
m21<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+balcony*bathrooms+zipcode
        +energy.eff+monument)

summary(m21)
#one of interaction significant, all varaibles signifcant
m31<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+balcony*energy.eff+bathrooms+zipcode+monument
        )
summary(m31)
#interaction terms are insignificant,all variables except zip
m41<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+balcony+energy.eff*bathrooms+zipcode+monument
        )
summary(m41)
#na interaction term, balcony is insignficant 
m51<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+balcony+bathrooms+
          +energy.eff*zipcode+monument)
summary(m51)
#insignficant interaction, balcony and energyy eff not signficant

```
```{r}

m6<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+balcony+zipcode+bathrooms+
          +energy.eff*monument)
summary(m6)
m7<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+balcony+zipcode+bathrooms*monument
          +energy.eff)
summary(m7)
m8<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+balcony*monument+zipcode+bathrooms+
          +energy.eff)
summary(m8)
m9<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+balcony+monument*zipcode+bathrooms+
          +energy.eff)
summary(m9)
```


#working with m2 best model from above 
```{r}
#model2
m21<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+balcony*bathrooms+zipcode
        +energy.eff+monument)
summary(m21)
#removing   balcony interaction as one of the variables are insignificant
m22<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+balcony+bathrooms+zipcode+energy.eff+monument)
summary(m22)
anova(m22,m21)
#reject take reduced model without interaction
#all balcony become insgnficant 
m23<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+bathrooms+zipcode+energy.eff+bathrooms
        +monument)
summary(m23)
anova(m23,m21)
#Reject, reduced model is better
#Dropping monuemnt varibales  as no catgories in it are signficant 
m24<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+bathrooms+zipcode+energy.eff)
anova(m24,m21)
# p value = 0.06 fail to reject h0, reduced model is better
summary(m24)
#The value of 2bathrooms was diffrent from our expication so we tested, interaction
interaction.plot(zipcode,bathrooms,retailvalue)
#No interaction
interaction.plot(zipcode,energy.eff,retailvalue)
#interaction presnt but when tested not signficant
#between zipcode and bathrooms and zipcode and energy.eff 
m25<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+bathrooms*zipcode+energy.eff)
#signficant
summary(m25)
par(mfrow=c(2,2))
plot(m25)
anova(m24,m25)
ols_coll_diag(m25)
#VIF shows no corealtion, but condition index does however the corealtion shown
#was tested if we could treat befor.

```

#Final Model
```{r}
m25<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+bathrooms*zipcode+energy.eff)
summary(m25)
par(mfrow=c(2,2))
plot(m25)
anova(m24,m25)
ols_coll_diag(m25)
```

#chacking hadis distance
```{r}
#Checking hadis distance and cooks
par(mfrow=c(1,2))
ols_plot_hadi(m25) #plotting hadi's 4points
ols_plot_cooksd_bar(m25)
#points are 2,15,21,50,52,59
```

#checking points using resdiuals
```{r}
#It is the scaled difference between the ith fitted value obtained from the full data 
#and the ith fitted value obtained by deleting the ith observation. DFFIT - difference 
#in fits, is used to identify influential data points. It quantifies the number of standard 
#deviations that the fitted value changes when the ith data point is omitted.
ols_plot_dffits(m25)
#Standardized Residual Chart. Chart for detecting outliers. Standardized residual 
#(internally studentized) is the residual divided by estimated standard deviatio
ols_plot_resid_stand(m25)
#Graph for detecting outliers.
ols_plot_resid_stud_fit(m25)
#common 2,15,59,35
#uncommon 21,50,52
#cooks wa hadi  2,15,21,50,52,59
```
#Testing removing points found 
```{r}
#Setting up dataset to be similar to the final data
dfi<-df[,c("retailvalue","lot.width","house.area","buildyear","bathrooms",
           "zipcode","energy.eff"), drop = FALSE]
dfi["zipcode"]<-ifelse(dfi["zipcode"]== '3528','3528','other')
dfi$bathrooms<-as.factor(dfi$bathrooms)
dfi$energy.eff<-as.factor(dfi$energy.eff)
dfi$zipcode<-as.factor(dfi$zipcode)
dfi1<-dfi[-c(selected_rows,2, 15, 35),]
m26<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+bathrooms*zipcode+energy.eff,data=dfi1)
summary(m26)
##

dfi2<-dfi[-c(selected_rows,21,50,52),]
m27<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+bathrooms*zipcode+energy.eff,data=dfi2)
summary(m27)
#
dfi3<-dfi[-c(selected_rows,59),]
m28<-lm(formula=retailvalue~I(lot.width^lambdalw) + I(house.area^lambdaha) +
          I(buildyear^lambdaby)+bathrooms*zipcode+energy.eff,data=dfi3)
summary(m28)
#
```

#Predictions 
```{r}
dft<-dfi[selected_rows,]
dft$bathrooms<-as.factor(dft$bathrooms)
dft$energy.eff<-as.factor(dft$energy.eff)
dft$zipcode<-as.factor(dft$zipcode)
predictions <- predict(m25, newdata = dft)
comparison_df <- cbind(dft, Predicted = predictions)
comparison_df
absolute_percentage_error<-abs(comparison_df$Predicted - comparison_df$retailvalue)/abs(comparison_df$Predicted)*100
# Display the results
print(paste(absolute_percentage_error))
```

#Predictions using the reduced models
```{r}
#m26,m27,m28
predictions2<- predict(m26, newdata = dft)
predictions3 <- predict(m27, newdata = dft)
predictions4 <- predict(m28, newdata = dft)
comparison_df <- cbind(dft, Predicted1=predictions,Predicted2 = predictions2,Predicted3 = predictions3, Predicted4 = predictions4)
comparison_df

absolute_percentage_error <- apply(comparison_df[, c("Predicted1", "Predicted2", "Predicted3", "Predicted4")], 2, function(x) {
  abs(x - comparison_df$retailvalue) / abs(x) * 100
})
# Add the errors to the comparison_df
comparison_df <- cbind(comparison_df, absolute_percentage_error)

# Display the results
print(comparison_df)
```

```
